---
title: "A Live View of COVID-19's Global Presence"
author: "Fan Wang"
date: "4/21/2020"
output: html_document
---

### Setup notebook
If you need to install these libraries, uncomment and run this cell:

```{r setup, include=TRUE, message=FALSE}
# install.packages("tidyr")
# install.packages("leaflet")
# install.packages("RCurl")
# install.packages("httr")
library(RCurl)
library(httr)
library(tidyverse)
library(leaflet)
knitr::opts_chunk$set(echo = TRUE)
```

Format the Date in the JHU dataframe as a date compatible with `as.Date()` in R:


```{r}
reformat.date <- function(data, col) {
    date_split <- str_split_fixed(data[[col]], "/", 3)
    storage.mode(date_split) <- "numeric"
    date_split <- as.data.frame(date_split)
    colnames(date_split) <- c("month", "day", "year")
    date_split <- date_split %>%
        mutate(
            month = str_pad(month, width = 2, pad = "0"),
            day = str_pad(day, width = 2, pad = "0"),
            year = paste0("20", year)
        )
    data[col] <- date_split %>% unite(col = "date", year, month, day, sep = "/", remove = TRUE, na.rm = FALSE)
    data[[col]] <- as.Date.character(data[[col]])
    data
}
```


### Loading in the Data

Data Source: COVID-19 Data Repository by the CSSE at Johns Hopkins University.

```{r, message=FALSE}
# world data
confirmed_table <-read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"),check.names=FALSE)

deaths_table <- read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"),check.names=FALSE)

recovered_table <- read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv"),check.names=FALSE)

pattern_date <- "^[[:digit:]]+/[[:digit:]]+/[[:digit:]]{2}$"
confirmed_table <- confirmed_table %>%
    pivot_longer(matches(pattern_date), names_to = "Date", values_to = "Cases") %>% 
    subset(!is.na(Cases))

confirmed_table <- reformat.date(confirmed_table, "Date")

deaths_table <- deaths_table %>%
    pivot_longer(matches(pattern_date), names_to = "Date", values_to = "Cases") %>% 
    subset(!is.na(Cases))

deaths_table <- reformat.date(deaths_table, "Date")

recovered_table <- recovered_table %>%
    pivot_longer(matches(pattern_date), names_to = "Date", values_to = "Cases") %>% 
    subset(!is.na(Cases))

recovered_table <- reformat.date(recovered_table, "Date")

# US data
confirmed_us_table <- read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"),check.names=FALSE)

deaths_us_table <- read.csv(text = getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"),check.names=FALSE)

confirmed_us_table <- confirmed_us_table %>%
    pivot_longer(matches(pattern_date), names_to = "Date", values_to = "Cases") %>% 
    subset(!is.na(Cases))

deaths_us_table <- deaths_us_table %>%
    pivot_longer(matches(pattern_date), names_to = "Date", values_to = "Cases") %>% 
    subset(!is.na(Cases))

confirmed_us_table <- reformat.date(confirmed_us_table, "Date")

deaths_us_table <- reformat.date(deaths_us_table, "Date")
```

### Setup data and parameters for plotting World Map

```{r, message=FALSE}
data_table <- merge(
    confirmed_table %>% rename(Confirmed = "Cases"),
    deaths_table %>% rename(Deaths = "Cases"),
    all = TRUE
) %>% select(-`Province/State`)
data_table <- merge(
    data_table,
    recovered_table %>% rename(Recovered = "Cases"),
    all = TRUE
) %>% select(-`Province/State`)
data_table <- data_table %>% mutate(
    Confirmed = ifelse(is.na(Confirmed), 0, Confirmed),
    Deaths = ifelse(is.na(Deaths), 0, Deaths),
    Recovered = ifelse(is.na(Recovered), 0, Recovered)
)
data_table <- data_table %>% mutate(Active = Confirmed - Deaths - Recovered)
plot_data <- data_table %>%
    subset(Date == max(Date)) %>%
    group_by(`Country/Region`) %>%
    summarise(
        Date = max(Date),
        Latitude = Lat[Confirmed == max(Confirmed)],
        Longitude = Long[Confirmed == max(Confirmed)],
        Confirmed = sum(Confirmed),
        Active = sum(Active),
        Deaths = sum(Deaths),
        Recovered = sum(Recovered)
    )

# Adjust the maximum multiplier to fix the maximum radius limit
radius_multiplier <- 6E5 / max(sqrt(plot_data$Confirmed), na.rm = TRUE)
```



### COVID-19 World Map at the latest time point

Click on the circles to open a popup with the latest COVID-19-related cases, deaths and recovered in country level.
(Note: the circle area is proportional to the confirmed cases.)

```{r, fig.width=8, fig.height=5, results='asis', warning=FALSE}
leaflet(plot_data) %>%
    addProviderTiles(providers$Esri.WorldImagery, group = "World Imagery") %>%
    addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
    addTiles() %>%
    addCircles(
        color = "red", lng = ~Longitude, lat = ~Latitude, weight = 1,
        radius = ~ sqrt(Confirmed) * radius_multiplier,
        popup = sprintf(
            paste(
                "<strong>%s</strong>",
                "<i>%s</i>",
                "Confirmed cases: %s",
                "Active cases: %s",
                "Death: %s",
                "Recovered: %s",
                sep = "</br>"
            ),
            format(plot_data$Date, "%a %b %d"),
            plot_data$`Country/Region`,
            format(plot_data$Confirmed, big.mark = ","),
            format(plot_data$Active, big.mark = ","),
            format(plot_data$Deaths, big.mark = ","),
            format(plot_data$Recovered, big.mark = ",")
        )
    )
```


### Setup data and parameters for plotting US Map

```{r, message=FALSE}
data_table <- merge(
    confirmed_us_table %>% rename(Confirmed = "Cases"),
    deaths_us_table %>% rename(Deaths = "Cases"),
    all = TRUE
) %>% select(-Province_State)
data_table <- data_table %>% mutate(
    Confirmed = ifelse(is.na(Confirmed), 0, Confirmed),
    Deaths = ifelse(is.na(Deaths), 0, Deaths)
)
data_table <- data_table %>%
    subset(Date == max(Date))

# Adjust the maximum multiplier to fix the maximum radius limit
radius_multiplier <- 2E5 / max(sqrt(data_table$Confirmed), na.rm = TRUE)
```

### COVID-19 U.S. Map at the latest time point

Click on the circles to open a popup with the latest COVID-19-related cases and deaths in county/region level.
(Note: the circle area is proportional to the confirmed cases.)

```{r, fig.width=8, fig.height=5, results='asis', warning=FALSE}
leaflet(data_table) %>%
    addProviderTiles(providers$Esri.WorldImagery, group = "World Imagery") %>%
    addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
    addTiles() %>%
    addCircles(
        color = "red", lng = ~Long_, lat = ~Lat, weight = 1,
        radius = ~ sqrt(Confirmed) * radius_multiplier,
        popup = sprintf(
            paste(
                "<strong>%s</strong>",
                "<i>%s</i>",
                "Confirmed cases: %s",
                "Death: %s",
                sep = "</br>"
            ),
            format(data_table$Date, "%a %b %d"),
            data_table$Admin2,
            format(data_table$Confirmed, big.mark = ","),
            format(data_table$Deaths, big.mark = ",")
        )
    )
```
